Spatio-angular fluorescence microscopy I. Basic theory

### 2018-11-05 papers/basic-theory/paper/paper.tex
Improved notation and aplanatic assumption.

% Talon's custom commands
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\argmin}{\arg\!\min}
\newcommand{\me}{\mathrm{e}}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\msf}[1]{\mathsf{#1}}
\providecommand{\mf}[1]{\mathfrak{#1}}
\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\ro}{\mathbf{\mathbf{r}}_o}
\providecommand{\so}{\mathbf{\hat{s}}_o}
\providecommand{\rb}{\mathbf{r}_b}
\providecommand{\rbm}[1]{r_b^{\text{m}}}
\providecommand{\rd}{\mathbf{r}_d}
\providecommand{\rdf}{\mathpzc{r}_d}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\bs}[1]{\boldsymbol{#1}}
\providecommand{\bv}{\bs{\nu}}
\providecommand{\bsh}[1]{\hat{\boldsymbol{#1}}}
\providecommand{\nan}{\left(\frac{\text{NA}}{n_o}\right)}
\providecommand{\lmsum}{\sum_{\ell=0}^\infty\sum_{m=-\ell}^{\ell}}
\providecommand{\intr}[1]{\int_{\mbb{R}^{#1}}}
\providecommand{\ints}[1]{\int_{\mbb{S}^{#1}}}
\providecommand{\tb}[1]{\textcolor{blue}{#1}}
% \DeclareFontFamily{OT1}{pzc}{}
% \DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
% \DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\newcommand{\eqname}[1]{\tag*{#1}}
\newcommand*\widefbox[1]{\fbox{\hspace{1em}#1\hspace{1em}}}

@@

The irradiance pattern measured by many fluorescence microscopes can be modeled
(at least approximately) as a magnifier with shift-invariant blur
\begin{align}
  g'(\rd') = \int_{\mbb{R}^2}d\ro\, h'(\rd' - m\ro)f(\ro), 
\end{align}
where $m$ is a magnification factor. To simplify our analysis we define a
demagnified detector coordinate $\rd = \rd'/m$ and a normalization factor that
corresponds to the total power incident on the detector plane due to a point
source $P_{\text{mono}} = \int_{\mbb{R}^2}d\mb{r}\,h'(m\mb{r})$. We use these scaling
factors to define the \textit{point spread function} as
\begin{align}
  h(\rd - \ro) = \frac{h'(m[\rd - \ro])}{P_{\text{mono}}},
\end{align}
and the \textit{scaled irradiance} as
\begin{align}
  g(\rd) = \frac{g'(m\rd)}{P_{\text{mono}}}.
\end{align}
With these definitions we can express the mapping between the object and the
data as a familiar convolution
\begin{align}
  g(\rd) = \int_{\mbb{R}^2}d\ro\, h(\rd - \ro)f(\ro),  \label{eq:lsi}
\end{align}
which will allow us to analyze this class of microscopes without carrying
magnification or normalization factors. We will mostly be using scaled
coordinates, so we have adopted a slightly unusual convention of using primes to
denote unscaled coordinates.

We have chosen to normalize the point spread function so that
\begin{align}
  \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}) = 1. \label{eq:norm}
\end{align}
The monopole point spread function corresponds to a measurable irradiance, so it
is always real-valued and positive.

@@

The mapping between the object and the data in linear shift-invariant
microscopes takes a particularly simple form in a spatial harmonic basis. If we
apply the Fourier convolution theorem to Eq. \ref{eq:lsi} we find that
\begin{align}
  G(\bs{\nu}) = H(\bs{\nu})F(\bs{\nu}),\label{eq:freq}
\end{align}
where we define the \textit{scaled irradiance spectrum} as
\begin{align}
  G(\bs{\nu}) = \int_{\mbb{R}^2}d\mb{r}\, g(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bs{\nu}),
\end{align}
the \textit{monopole transfer function} as
\begin{align}
  H(\bs{\nu}) = \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bs{\nu}),\label{eq:otf}
\end{align}
and the \textit{monopole spectrum} as
\begin{align}
    F(\bs{\nu}) = \int_{\mbb{R}^2}d\mb{r}\, f(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bs{\nu}).
\end{align}
The monopole point spread function is normalized and real-valued, so we know
that the monopole transfer function is normalized $H(0) = 1$ and
conjugate-symmetric $H(-\bs{\nu}) = H^*(\bs{\nu})$.

@@

\subsubsection{Monopole coherent transfer functions}
Although the Fourier transform can be used to calculate the monopole transfer
function directly from the monopole point spread function, there is a well-known
alternative that exploits coherent transfer functions. The key idea is that the
monopole point spread function can always be written as the absolute square of a
scalar-valued \textit{monopole coherent spread function} defined as
\begin{align}
  h(\mb{r}) = |c(\mb{r})|^2. \label{eq:absquarescalar}
\end{align}
Physically, the monopole coherent spread function is the scalar-valued field on
the detector with appropriate scaling.

Since the monopole transfer function is given by the two-dimensional Fourier
transform of the monopole point spread function
\begin{align}
  H(\bs{\nu}) = \int_{\mbb{R}^2}d\mb{r}\,h(\mb{r})\,\text{exp}[-2\pi i\mb{r}\cdot\bs{\nu}],
\end{align}
we can plug in Eq. \ref{eq:absquarescalar} and use the autocorrelation theorem to
rewrite the monopole transfer function as
\begin{align}
  H(\bs{\nu}) = \int_{\mbb{R}^2}d\bs{\tau}\, C(\bs{\tau})C^*(\bs{\tau} - \bs{\nu}), 
\end{align}
where $z^*$ denotes the complex conjugate of $z$, and we have introduced the
\textit{monopole coherent transfer function} as the two-dimensional Fourier
transform of the orientation-dependent coherent spread function
\begin{align}
  C(\bs{\tau}) = \int_{\mbb{R}^2}d\mb{r}\, c(\mb{r})\,\text{exp}[-2\pi i\mb{r}\cdot\bs{\tau}].
\end{align}
Physically, the monopole coherent transfer function is the scalar-valued field
in a Fourier plane of the detector (often called the pupil or aperture plane)
with appropriate scaling. This provides a valuable shortcut for analyzing
arbitrary microscopes since it is usually straightforward to calculate the field
in the pupil. A typical approach for calculating the transfer functions is to
(1) calculate the field in the pupil, (2) scale the field to find the coherent
monopole transfer function, then (3) use the relationships in Figure
\ref{fig:monopole-transfer-functions} to calculate the other transfer functions.

@@

In a delta function basis the object can be represented by a function
$f(\ro, \so)$ called the \textit{dipole density}---the number of dipoles at
position $\ro{}$ per unit area oriented along $\so{}$ per unit solid angle.
Similar to the monopole case, we model the mapping between the object and the
irradiance in a delta function basis as an integral transform
\begin{align}
  g'(\rd') = \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\ro\, h'(\rd', \ro, \so)f(\ro, \so). 
\end{align}
where $h'(\rd', \ro, \so)$ is the irradiance at position $\rd'$ created by a
point source at $\ro$ with orientation $\so$. Once again, many fluorescence
microscopes can be modeled as magnifiers with shift-invariant blur
\begin{align}
  g'(\rd') = \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\ro\, h'(\rd' - m\ro, \so)f(\ro, \so). 
\end{align}

We define the same demagnified detector coordinate $\rd = \rd'/m$ and a new
normalization factor that corresponds to the total power incident on the
detector due to a spatial point source with an angularly uniform distribution of
dipoles
$P_\text{dip} = \int_{\mbb{S}^2}d\mh{s}\int_{\mbb{R}^2}d\mb{r}\, h'(m\mb{r},
\mh{s})$. We use these scaling factors to define the
\textit{dipole point spread function} as
\begin{align}
  h(\rd - \ro, \so) = \frac{h'(m[\rd - \ro], \so)}{P_\text{dip}}, 
\end{align}
and the \textit{scaled irradiance} as 
\begin{align}
  g(\rd) = \frac{g'(m\rd)}{P_\text{dip}}. 
\end{align}
With these definitions we can express the mapping between the object and the
data as
\begin{align}
g(\rd{}) = \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^2}d\ro{}\, h(\rd{} -\ro{}, \so{})f(\ro, \so). \label{eq:odpsf}
\end{align}

Similar to the monopole case, we have chosen to normalize the dipole point
spread function so that
\begin{align}
  \int_{\mbb{S}^2}d\mh{s}\int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}, \mh{s}) = 1. 
\end{align}
The dipole point spread function is a measurable quantity, so it is real-valued
and positive.

\subsubsection{Spatial dipole transfer function}
We can make our first change of basis by applying the Fourier-convolution
theorem to Eq. \ref{eq:odpsf} which yields
\begin{align}
G(\bs{\nu}) = \int_{\mbb{S}^2}d\mh{s}\, H(\bs{\nu}, \mh{s})F(\bs{\nu}, \mh{s}) \label{eq:odotf},
\end{align}
where we define the \textit{spatial dipole transfer function} as
  \begin{align}
  H(\bs{\nu}, \mh{s}) &= \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}, \mh{s})\, \text{exp}(-2\pi i\mb{r}\cdot\bs{\nu}),
  \end{align}
  and the \textit{spatial dipole spectrum} as
  \begin{align}
  F(\bs{\nu}, \mh{s}) &= \int_{\mbb{R}^2}d\mb{r}\, f(\mb{r}, \mh{s})\, \text{exp}(-2\pi i\mb{r}\cdot\bs{\nu}). 
  \end{align}
  Since the dipole point spread function is normalized and real, we know that
  the spatial dipole transfer function is normalized
  $\int_{\mbb{S}^2}d\mh{s}\, H(0, \mh{s}) = 1$ and conjugate-symmetric
  $H(-\bs{\nu}, \mh{s}) = H^*(\bs{\nu}, \mh{s})$.

@@

\subsubsection{Angular dipole transfer function}
The spherical harmonics are another set of convenient basis functions. We can
change basis from spherical delta functions to spherical harmonics by applying
the generalized Plancharel theorem for spherical functions
\begin{align}
  \ints{2}d\mh{s}{}\, p(\mh{s})q(\mh{s}) = \lmsum P_\ell^m Q_\ell^m, \label{eq:plan}
\end{align}
where $p(\mh{s})$ and $q(\mh{s})$ are arbitrary functions on the sphere, $P_\ell^m$ and $Q_\ell^m$ are their spherical Fourier transforms defined by
\begin{align}
  P_\ell^m = \int_{\mbb{S}^2}d\mh{s}\, p(\mh{s})Y_\ell^{m*}(\mh{s}),
\end{align}
and $Y_{\ell}^m(\mh{s})$ are the spherical harmonic functions defined in
Appendix \ref{sec:sph}. Eq. \ref{eq:plan} expresses the fact that scalar
products are invariant under a change of basis (see Eq. 3.78 of
\cite{barrett2004}). The left-hand side of Eq. \ref{eq:plan} is the scalar
product of $\mbb{L}_2(\mbb{S}^2)$ functions in a delta function basis and the
right hand side is the scalar product of $\mbb{L}_2(\mbb{S}^2)$ functions in a
spherical harmonic function basis. Applying Eq. \ref{eq:plan} to Eq.
\ref{eq:odpsf} yields
\begin{align}
  g(\rd) = \lmsum H_\ell^m(\rd - \ro)F_\ell^m(\ro) \label{eq:atf-form}
\end{align}
where we have defined the \textit{angular dipole transfer function} as
\begin{align}
  H_\ell^m(\mb{r}) = \int_{\mbb{S}^2}d\mh{s}\, h(\mb{r}, \mh{s})Y_{\ell}^{m*}(\mh{s}),\label{eq:atf-prep} 
\end{align}
and the \textit{angular dipole spectrum} as
\begin{align}
  F_\ell^m(\mb{r}) = \int_{\mbb{S}^2}d\mh{s}\, f(\mb{r}, \mh{s})Y_{\ell}^{m*}(\mh{s}).
\end{align}
Since the dipole point spread function is normalized and real-valued, we know
that the angular dipole transfer function is normalized
$\int_{\mbb{R}^2}d\mb{r}\, H_0^0(\mb{r}) = 1$ and conjugate symmetric
$H_\ell^{-m}(\mb{r}) = (-1)^mH_\ell^{m*}(\mb{r})$.

@@

\subsubsection{Spatio-angular dipole transfer function}
We can arrive at our final basis in two ways: by applying the generalized
Plancharel theorem for spherical functions to Eq. \ref{eq:odotf} or by applying
the Fourier convolution theorem to Eq. \ref{eq:atf-form}. We follow the
second path and find that
\begin{align}
G(\bs{\nu}) = \lmsum \msf{H}_\ell^m(\bs{\nu})\msf{F}_\ell^m(\bs{\nu}) \label{eq:saft},
\end{align}
where we have defined the \textit{spatio-angular dipole transfer function} as
  \begin{align}
  \msf{H}_\ell^m(\bs{\nu}) &= \int_{\mbb{S}^2}d\mh{s}\, H(\bs{\nu}, \mh{s})Y_\ell^{m*}(\mh{s}),
  \end{align}
  and the \textit{spatio-angular dipole spectrum} as
  \begin{align}
  \msf{F}_\ell^m(\bs{\nu}) &= \int_{\mbb{S}^2}d\mh{s}\, F(\bs{\nu}, \mh{s})Y_\ell^{m*}(\mh{s}).
  \end{align}
  Since the dipole point spread function is normalized and real-valued, we know
  that the spatio-angular dipole transfer function is normalized
  $\msf{H}_0^0(0) = 1$ and conjugate symmetric
  $\msf{H}_\ell^{-m}(-\bs{\nu}) = (-1)^m\msf{H}_\ell^{m*}(\bs{\nu})$.

@@

\subsubsection{Dipole coherent transfer functions}
Similar to the monopole case, there is efficient way to calculate the transfer
functions using coherent transfer functions. The dipole point spread function
can always be written as the absolute square of a vector-valued function called
the \textit{dipole coherent spread function}
\begin{align}
  h(\mb{r}, \mh{s}) = |\mb{c}(\mb{r}, \mh{s})|^2. \label{eq:absquare2}
\end{align}
Physically, the dipole coherent spread function is an appropriately scaled
vector-valued electric field on the detector. We need a vector-valued coherent
transfer function since the polarization of the field plays a significant role
in dipole imaging.

Since the spatial dipole transfer function is given by the two-dimensional
Fourier transform of the dipole point spread function
\begin{align}
  H(\bs{\nu}) = \int_{\mbb{R}^2}d\mb{r}\,h(\mb{r}, \mh{s})\,\text{exp}[-2\pi i\mb{r}\cdot\bs{\nu}],
\end{align}
we can plug in Eq. \ref{eq:absquare2} and use the autocorrelation theorem to
rewrite the spatial dipole transfer function as 
\begin{align}
  H(\bs{\nu}) = \int_{\mbb{R}^2}d\bs{\tau}\,\mb{C}(\bs{\tau}, \mh{s})\mb{C}^\dagger(\bs{\tau} - \bs{\nu}, \mh{s}), 
\end{align}
where we have introduced the \textit{dipole coherent transfer function}
$\mb{C}(\bs{\tau}, \mh{s})$ as the two-dimensional Fourier transform of the
dipole coherent spread function
\begin{align}
  \mb{C}(\bs{\tau}, \mh{s}) = \int_{\mbb{R}^2}d\mb{r}\, \mb{c}(\mb{r}, \mh{s})\,\text{exp}[-2\pi i\mb{r}\cdot\bs{\tau}].
\end{align}
Physically, the dipole coherent transfer function corresponds to the
vector-valued electric field created by a dipole with oriented along $\mh{s}$ in
a Fourier plane of the detector with appropriate scaling. As with the monopole
case we can calculate the dipole-orientation-dependent fields in the pupil
plane, scale them appropriately to find the dipole coherent transfer function,
the use the relationships in Figure \ref{fig:transfer-functions} to calculate
the other transfer functions. Note that we could define coherent equivalents
for all four incoherent transfer functions, but we restrict ourselves to these
two since they are the most useful for our calculations.

@@

\subsection{Monopole point spread function of a paraxial epi-fluorescence microscope}
In this section we derive the form of the monopole point spread function
$h(\rd - \ro)$. First, we place a monopole emitter at the focal point of an
objective lens and calculate the field $U_{\text{bfp}}(\rb)$ created in the back
focal plane. We can model a paraxial thin lens as a quadratic phase element, so
it converts an incident spherical wave to a plane wave in the back focal plane
with a sharp cutoff at the exit pupil of the lens
\begin{align}
  U_{\text{bfp}}(\mb{r}_b) \propto \Pi\left(\frac{|\mb{r}_b|}{2\alpha}\right) \equiv
  \begin{cases}
    1\quad \text{if}\quad |\mb{r}_b| < \alpha,\\
    0\quad \text{else},
  \end{cases}
\end{align}
where $\rb$ is the two-dimensional coordinate in the back focal plane and
$\alpha$ is the collection half angle---see Figure XXX.

>>>>

% Talon's custom commands
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\argmin}{\arg\!\min}
\newcommand{\me}{\mathrm{e}}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\msf}[1]{\mathsf{#1}}
\providecommand{\mf}[1]{\mathfrak{#1}}
\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\ro}{\mathbf{\mathbf{r}}_o}
\providecommand{\so}{\mathbf{\hat{s}}_o}
\providecommand{\rp}{\mathbf{r}_p}
\providecommand{\rbm}[1]{r_b^{\text{m}}}
\providecommand{\rd}{\mathbf{r}_d}
\providecommand{\rdf}{\mathpzc{r}_d}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\bs}[1]{\boldsymbol{#1}}
\providecommand{\bv}{\bs{\nu}}
\providecommand{\bvp}{\bs{\nu}_p}
\providecommand{\bsh}[1]{\hat{\boldsymbol{#1}}}
\providecommand{\nan}{\left(\frac{\text{NA}}{n_o}\right)}
\providecommand{\lmsum}{\sum_{\ell=0}^\infty\sum_{m=-\ell}^{\ell}}
\providecommand{\intr}[1]{\int_{\mbb{R}^{#1}}}
\providecommand{\ints}[1]{\int_{\mbb{S}^{#1}}}
\providecommand{\tb}[1]{\textcolor{blue}{#1}}
% \DeclareFontFamily{OT1}{pzc}{}
% \DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
% \DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\newcommand{\eqname}[1]{\tag*{#1}}
\newcommand*\widefbox[1]{\fbox{\hspace{1em}#1\hspace{1em}}}

@@

Next, we assume that the optical system is \textit{aplanatic}---Abbe's sine
condition is satisfied and on-axis points are imaged without aberration. Abbe's
sine condition guarantees that off-axis points are imaged without spherical
aberration or coma \cite{mansuripur2009}, so the imaging system can be modeled
(at least approximately) as a magnifier with shift-invariant blur
\begin{align}
  g'(\rd') = \int_{\mbb{R}^2}d\ro\, h'(\rd' - m\ro)f(\ro), \label{eq:nonconv}
\end{align}
where $m$ is a magnification factor.

We can simplify our analysis by changing coordinates and writing Eq.
\ref{eq:nonconv} as a convolution \cite{barrett2004}. We define a demagnified
detector coordinate $\rd = \rd'/m$ and a normalization factor that corresponds
to the total power incident on the detector plane due to a point source
$P_{\text{mono}} = \int_{\mbb{R}^2}d\mb{r}\,h'(m\mb{r})$ where
$\mb{r} = \rd - \ro$. We use these scaling factors to define the
\textit{monopole point spread function} as
\begin{align}
  h(\rd - \ro) = \frac{h'(m[\rd - \ro])}{P_{\text{mono}}},
\end{align}
and the \textit{scaled irradiance} as
\begin{align}
  g(\rd) = \frac{g'(m\rd)}{P_{\text{mono}}}.
\end{align}
With these definitions we can express the mapping between the object and the
data as a familiar convolution
\begin{align}
  g(\rd) = \int_{\mbb{R}^2}d\ro\, h(\rd - \ro)f(\ro).  \label{eq:lsi}
\end{align}
We will mostly be using scaled coordinates, so we have adopted a slightly
unusual convention of using primes to denote unscaled coordinates.

We have chosen to normalize the monopole point spread function so that
\begin{align}
  \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}) = 1. \label{eq:norm}
\end{align}
The monopole point spread function corresponds to a
measurable irradiance, so it is always real-valued and positive.

@@

The mapping between the object and the data in linear shift-invariant
microscopes takes a particularly simple form in a spatial harmonic basis. If we
apply the Fourier convolution theorem to Eq. \ref{eq:lsi} we find that
\begin{align}
  G(\bv) = H(\bv)F(\bv),\label{eq:freq}
\end{align}
where we define the \textit{scaled irradiance spectrum} as
\begin{align}
  G(\bv) = \int_{\mbb{R}^2}d\mb{r}\, g(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bv),
\end{align}
the \textit{monopole transfer function} as
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bv),\label{eq:otf}
\end{align}
and the \textit{monopole spectrum} as
\begin{align}
    F(\bv) = \int_{\mbb{R}^2}d\mb{r}\, f(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bv).
\end{align}
The monopole point spread function is normalized and real-valued, so we know
that the monopole transfer function is normalized $H(0) = 1$ and
conjugate-symmetric $H(-\bv) = H^*(\bv)$.

@@

\subsubsection{Monopole coherent transfer functions}
Although the Fourier transform can be used to calculate the monopole transfer
function directly from the monopole point spread function, there is a well-known
alternative that exploits coherent transfer functions. The key idea is that the
monopole point spread function can always be written as the absolute square of a
scalar-valued \textit{monopole coherent spread function} defined as
\begin{align}
  |c(\rd - \ro)|^2 = h(\rd - \ro). \label{eq:absquarescalar}
\end{align}
Physically, the monopole coherent spread function corresponds to the
scalar-valued field on the detector with appropriate scaling.

Since the monopole transfer function is given by the two-dimensional Fourier
transform of the monopole point spread function
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\mb{r}\,h(\mb{r})\,\text{exp}[-2\pi i\mb{r}\cdot\bv],
\end{align}
we can plug in Eq. \ref{eq:absquarescalar} and use the autocorrelation theorem to
rewrite the monopole transfer function as
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\bvp\, C(\bvp)C^*(\bvp - \bv), 
\end{align}
where $z^*$ denotes the complex conjugate of $z$, and we have introduced the
\textit{monopole coherent transfer function} as the two-dimensional Fourier
transform of the monopole coherent spread function
\begin{align}
  C(\bvp) = \int_{\mbb{R}^2}d\mb{r}\, c(\mb{r})\,\text{exp}[-2\pi i\mb{r}\cdot\bvp].
\end{align}
Physically, the monopole coherent transfer function corresponds to the
scalar-valued field in a Fourier plane of the detector with appropriate scaling.
We have chosen the symbol $\bvp$ as the coordinate for the coherent
transfer function because the Fourier plane of the detector is often called the
\textit{pupil plane}.

The coherent transfer function provides a valuable shortcut for analyzing
microscopes since it is usually straightforward to calculate the field in the
aperture plane. A typical approach for calculating the transfer functions is to
(1) calculate the field in the aperture plane, (2) scale the field to find the
coherent monopole transfer function, then (3) use the relationships in Figure
\ref{fig:monopole-transfer-functions} to calculate the other transfer functions.

@@

In a delta function basis the object can be represented by a function
$f(\ro, \so)$ called the \textit{dipole density}---the number of dipoles at
position $\ro{}$ per unit area oriented along $\so{}$ per unit solid angle.
Similar to the monopole case, we model the mapping between the object and the
irradiance in a delta function basis as an integral transform
\begin{align}
  g'(\rd') = \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\ro\, h'(\rd', \ro, \so)f(\ro, \so). 
\end{align}
where $h'(\rd', \ro, \so)$ is the irradiance at position $\rd'$ created by a
point source at $\ro$ with orientation $\so$. If the imaging system is
linear and aplanatic we can write the integral transform as
\begin{align}
  g'(\rd') = \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\ro\, h'(\rd' - m\ro, \so)f(\ro, \so). 
\end{align}

We define the same demagnified detector coordinate $\rd = \rd'/m$ and a new
normalization factor that corresponds to the total power incident on the
detector due to a spatial point source with an angularly uniform distribution of
dipoles
$P_\text{dip} = \int_{\mbb{S}^2}d\mh{s}\int_{\mbb{R}^2}d\mb{r}\, h'(m\mb{r},
\so)$. We use these scaling factors to define the
\textit{dipole point spread function} as
\begin{align}
  h(\rd - \ro, \so) = \frac{h'(m[\rd - \ro], \so)}{P_\text{dip}}, 
\end{align}
and the \textit{scaled irradiance} as 
\begin{align}
  g(\rd) = \frac{g'(m\rd)}{P_\text{dip}}. 
\end{align}
With these definitions we can express the mapping between the object and the
data as
\begin{align}
g(\rd{}) = \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^2}d\ro{}\, h(\rd{} -\ro{}, \so{})f(\ro, \so). \label{eq:odpsf}
\end{align}

Similar to the monopole case, we have chosen to normalize the dipole point
spread function so that
\begin{align}
  \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}, \so) = 1. 
\end{align}
The dipole point spread function is a measurable quantity, so it is real-valued
and positive.

\subsubsection{Spatial dipole transfer function}
We can make our first change of basis by applying the Fourier-convolution
theorem to Eq. \ref{eq:odpsf} which yields
\begin{align}
G(\bv) = \int_{\mbb{S}^2}d\so\, H(\bv, \so)F(\bv, \so) \label{eq:odotf},
\end{align}
where we define the \textit{spatial dipole transfer function} as
  \begin{align}
  H(\bv, \so) &= \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}, \so)\, \text{exp}(-2\pi i\mb{r}\cdot\bv),
  \end{align}
  and the \textit{spatial dipole spectrum} as
  \begin{align}
  F(\bv, \so) &= \int_{\mbb{R}^2}d\mb{r}\, f(\mb{r}, \so)\, \text{exp}(-2\pi i\mb{r}\cdot\bv). 
  \end{align}
  Since the dipole point spread function is normalized and real, we know that
  the spatial dipole transfer function is normalized
  $\int_{\mbb{S}^2}d\mh{s}\, H(0, \so) = 1$ and conjugate-symmetric
  $H(-\bv, \so) = H^*(\bv, \so)$.

@@

\subsubsection{Angular dipole transfer function}
The spherical harmonics are another set of convenient basis functions. We can
change basis from spherical delta functions to spherical harmonics by applying
the generalized Plancharel theorem for spherical functions
\begin{align}
  \ints{2}d\mh{s}\, p(\mh{s})q(\mh{s}) = \lmsum P_\ell^m Q_\ell^m, \label{eq:plan}
\end{align}
where $p(\mh{s})$ and $q(\mh{s})$ are arbitrary functions on the sphere, $P_\ell^m$ and $Q_\ell^m$ are their spherical Fourier transforms defined by
\begin{align}
  P_\ell^m = \int_{\mbb{S}^2}d\mh{s}\, p(\mh{s})Y_\ell^{m*}(\mh{s}),
\end{align}
and $Y_{\ell}^m(\mh{s})$ are the spherical harmonic functions defined in
Appendix \ref{sec:sph}. Eq. \ref{eq:plan} expresses the fact that scalar
products are invariant under a change of basis (see Eq. 3.78 of
\cite{barrett2004}). The left-hand side of Eq. \ref{eq:plan} is the scalar
product of $\mbb{L}_2(\mbb{S}^2)$ functions in a delta function basis and the
right-hand side is the scalar product of $\mbb{L}_2(\mbb{S}^2)$ functions in a
spherical harmonic function basis. Applying Eq. \ref{eq:plan} to Eq.
\ref{eq:odpsf} yields
\begin{align}
  g(\rd) = \lmsum H_\ell^m(\rd - \ro)F_\ell^m(\ro) \label{eq:atf-form}
\end{align}
where we have defined the \textit{angular dipole transfer function} as
\begin{align}
  H_\ell^m(\rd - \ro) = \int_{\mbb{S}^2}d\so\, h(\rd - \ro, \so)Y_{\ell}^{m*}(\so),\label{eq:atf-prep} 
\end{align}
and the \textit{angular dipole spectrum} as
\begin{align}
  F_\ell^m(\ro) = \int_{\mbb{S}^2}d\so\, f(\ro, \so)Y_{\ell}^{m*}(\so).
\end{align}
Since the dipole point spread function is normalized and real-valued, we know
that the angular dipole transfer function is normalized
$\int_{\mbb{R}^2}d\mb{r}\, H_0^0(\mb{r}) = 1$ and conjugate symmetric
$H_\ell^{-m}(\mb{r}) = (-1)^mH_\ell^{m*}(\mb{r})$.

@@

\subsubsection{Spatio-angular dipole transfer function}
We can arrive at our final basis in two ways: by applying the generalized
Plancharel theorem for spherical functions to Eq. \ref{eq:odotf} or by applying
the Fourier convolution theorem to Eq. \ref{eq:atf-form}. We follow the
second path and find that
\begin{align}
G(\bv) = \lmsum \msf{H}_\ell^m(\bv)\msf{F}_\ell^m(\bv) \label{eq:saft},
\end{align}
where we have defined the \textit{spatio-angular dipole transfer function} as
  \begin{align}
  \msf{H}_\ell^m(\bv) &= \int_{\mbb{S}^2}d\so\, H(\bv, \so)Y_\ell^{m*}(\so),
  \end{align}
  and the \textit{spatio-angular dipole spectrum} as
  \begin{align}
  \msf{F}_\ell^m(\bv) &= \int_{\mbb{S}^2}d\so\, F(\bv, \so)Y_\ell^{m*}(\so).
  \end{align}
  Since the dipole point spread function is normalized and real-valued, we know
  that the spatio-angular dipole transfer function is normalized
  $\msf{H}_0^0(0) = 1$ and conjugate symmetric
  $\msf{H}_\ell^{-m}(-\bv) = (-1)^m\msf{H}_\ell^{m*}(\bv)$.

@@

\subsubsection{Dipole coherent transfer functions}
Similar to the monopole case, there is efficient way to calculate the transfer
functions using coherent transfer functions. The dipole point spread function
can always be written as the absolute square of a vector-valued function called
the \textit{dipole coherent spread function}
\begin{align}
  |\mb{c}(\rd - \ro, \so)|^2 = h(\rd - \ro, \so). \label{eq:absquare2}
\end{align}
Physically, the dipole coherent spread function corresponds to the vector-valued
electric field on the detector with appropriate scaling. We need a vector-valued
coherent transfer function since the polarization of the field plays a
significant role in dipole imaging, so the point spread function cannot be
written as an absolute square of a scalar-valued function.

Since the spatial dipole transfer function is given by the two-dimensional
Fourier transform of the dipole point spread function
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\mb{r}\,h(\mb{r}, \so)\,\text{exp}[-2\pi i\mb{r}\cdot\bv],
\end{align}
we can plug in Eq. \ref{eq:absquare2} and use the autocorrelation theorem to
rewrite the spatial dipole transfer function as 
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\bvp\,\mb{C}(\bvp, \so)\mb{C}^\dagger(\bvp - \bv, \so), 
\end{align}
where we have introduced the \textit{dipole coherent transfer function}
$\mb{C}(\bvp, \so)$ as the two-dimensional Fourier transform of the dipole
coherent spread function
\begin{align}
  \mb{C}(\bvp, \so) = \int_{\mbb{R}^2}d\mb{r}\, \mb{c}(\mb{r}, \so)\,\text{exp}[-2\pi i\mb{r}\cdot\bvp].
\end{align}
Physically, the dipole coherent transfer function corresponds to the
vector-valued electric field created by a dipole oriented along $\so$ in a
Fourier plane of the detector with appropriate scaling. Similar to the monopole
case, we can calculate the orientation-dependent fields in the aperture plane,
scale appropriately to find the dipole coherent transfer function, then use the
relationships in Figure \ref{fig:transfer-functions} to calculate the other
transfer functions. Note that we could define coherent equivalents for all four
incoherent transfer functions, but we restrict ourselves to these two since they
are the most useful for our calculations.

@@

\subsection{Monopole point spread function of a paraxial epi-fluorescence microscope}
In this section we derive the form of the monopole point spread function
$h(\rd - \ro)$. First, we place a monopole emitter at the focal point of an
objective lens and calculate the field $U_{\text{bfp}}(\rp)$ created in the back
focal plane. We can model a paraxial thin lens as a quadratic phase element, so
it converts an incident spherical wave to a plane wave in the back focal plane
with a sharp cutoff at the exit pupil of the lens
\begin{align}
  U_{\text{bfp}}(\mb{r}_b) \propto \Pi\left(\frac{|\mb{r}_b|}{2\alpha}\right) \equiv
  \begin{cases}
    1\quad \text{if}\quad |\mb{r}_b| < \alpha,\\
    0\quad \text{else},
  \end{cases}
\end{align}
where $\rp$ is the two-dimensional coordinate in the back focal plane and
$\alpha$ is the collection half angle---see Figure XXX.
